---
title: "Terrorists Attacks"
author: "Ahmed Battah"
date: "March 7, 2018"
output: html_document
---

```{r setup, include = FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = FALSE}

library(tidyverse)
library(ggplot2)
library(leaflet)

terrorists_attack <- read_csv('terrorism_attack')

terrorists_attack <-
  terrorists_attack %>% select(
    -c(
      approxdate,
      extended,
      country,
      country_txt,
      region,
      region_txt,
      attacktype3,
      attacktype3_txt,
      guncertain3,
      claimmode2,
      claimmode2_txt,
      claimmode3,
      claimmode3_txt,
      compclaim,
      weaptype4,
      weapsubtype4_txt,
      nhostkid,
      nhostkidus,
      nhours,
      ndays,
      divert,
      kidhijcountry,
      ransomamt,
      ransomamtus,
      ransompaid,
      ransompaidus,
      ransomnote,
      hostkidoutcome,
      hostkidoutcome_txt,
      nreleased
    )
  )

terrorists_attack <- terrorists_attack %>%
  mutate(decade = ifelse(iyear < 1980, "1970s",
                         ifelse(
                           iyear < 1990, "1980s",
                           ifelse(iyear < 2000, "1990s",
                                  ifelse(iyear < 2010, "2000s", "2010s"))
                         )))
terrorists_attack$decade <- factor(terrorists_attack$decade,
                                   levels = c("1970s", "1980s", "1990s",
                                              "2000s", "2010s"))


```

### Overview

The database is maintained by researchers at the National Consortium for the Study of Terrorism and Responses to Terrorism (START), headquartered at the University of Maryland. [More Information](http://start.umd.edu/gtd/).

Geography: Bahrain  
Time period: 1970-2016

#### Preview of data
```{r eval = TRUE, echo = FALSE, results = 'asis'}
kable(terrorists_attack[1:10, c("city", "latitude", "longitude", "attacktype1_txt", "nkill", "nwound")])
```

* Data dimension: **`r nrow(terrorists_attack)`** x **`r ncol(terrorists_attack)`**
 
 ***
 
### Data Visualization
  
#### Number of Killed/Wounded by year
```{r eval = TRUE, echo = FALSE}
terrorists_attack %>% 
  filter(nkill > 0 | nwound > 0) %>% 
  group_by(iyear) %>% 
  summarise(Killed = sum(nkill), Wounded = sum(nwound)) %>% 
  ungroup() %>% 
ggplot(aes(x = iyear)) +
  geom_line(aes(y = Killed, colour = "Killed")) + 
  geom_point(aes(y = Killed, colour = "Killed")) + 
  geom_line(aes(y = Wounded, colour = "Wounded")) + 
  geom_point(aes(y = Wounded, colour = "Wounded")) + 
  scale_x_continuous(breaks = terrorists_attack$iyear) +
  labs(title = "", x = "Year", y = "Count", colour = "") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45,
                       hjust = 1,
                       vjust = 1))
```

#### Number Attacks per decade
```{r eval = TRUE, echo = FALSE}
ggplot(terrorists_attack, aes(x = decade, fill = decade)) +
  geom_bar() +
  labs(title = "", x = "Decade", y = "Count", fill = "") +
  geom_text(stat = "count", aes(label = ..count..), 
            vjust = -0.2, size = 3) +
  theme_bw()
```


