---
title: "Terrorists Attacks"
author: "Ahmed Battah"
date: "March 10, 2018"
output: html_document
---

```{r setup, include = FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = FALSE}

library(tidyverse)
library(ggplot2)
library(leaflet)
library(kableExtra)
library(RColorBrewer)

terrorists_attack <- read_csv('terrorism_attack')

terrorists_attack <-
  terrorists_attack %>% select(
    -c(
      approxdate,
      extended,
      country,
      country_txt,
      region,
      region_txt,
      attacktype3,
      attacktype3_txt,
      guncertain3,
      claimmode2,
      claimmode2_txt,
      claimmode3,
      claimmode3_txt,
      compclaim,
      weaptype4,
      weapsubtype4_txt,
      nhostkid,
      nhostkidus,
      nhours,
      ndays,
      divert,
      kidhijcountry,
      ransomamt,
      ransomamtus,
      ransompaid,
      ransompaidus,
      ransomnote,
      hostkidoutcome,
      hostkidoutcome_txt,
      nreleased
    )
  )

# Fix city names
terrorists_attack$city <- terrorists_attack$city %>% 
  recode("Wadiyan" = "Wadyan",
         "Sitrah" = "Sitra",
         "Sahla" = "Sehla",
         "Quraiya" = "Qurayyah",
         "Karzakkan" = "Karzakan",
         "Duraz" = "Diraz",
         "Daih" = "Diah",
         "Demistan" = "Damistan",
         "Al-Akar" = "Eker",
         "Akr" = "Eker")

terrorists_attack <- terrorists_attack %>%
  mutate(decade = ifelse(iyear < 1980, "1970s",
                         ifelse(
                           iyear < 1990, "1980s",
                           ifelse(iyear < 2000, "1990s",
                                  ifelse(iyear < 2010, "2000s", "2010s"))
                         )))
terrorists_attack$decade <- factor(terrorists_attack$decade,
                                   levels = c("1970s", "1980s", "1990s",
                                              "2000s", "2010s"))


```

### Overview

The database is maintained by researchers at the National Consortium for the Study of Terrorism and Responses to Terrorism (START), headquartered at the University of Maryland. [More Information](http://start.umd.edu/gtd/).

Geography: Bahrain  
Time period: 1970-2016

<br><br>

#### Preview of data
```{r eval = TRUE, echo = FALSE, results = 'asis'}
kable(terrorists_attack[1:10, c("provstate", "city", "attacktype1_txt", "target1", "nkill", "nwound")], col.names = c("Governate", "City","Attack type","Target","Killed","Wounded"), "html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

* Data dimension: **`r nrow(terrorists_attack)`** x **`r ncol(terrorists_attack)`**
 
 ***

<br>

### Data Visualization

#### Number Attacks per decade
```{r eval = TRUE, echo = FALSE}
ggplot(terrorists_attack, aes(x = decade, fill = decade)) +
  geom_bar() +
  labs(title = "", x = "", y = "", fill = "Decade") +
  geom_text(stat = "count", aes(label = ..count..), 
            vjust = -0.2, size = 3) +
  theme_bw()
```

<br> 
  
#### Number of Killed/Wounded by year
```{r eval = TRUE, echo = FALSE}
terrorists_attack %>% 
  filter(nkill > 0 | nwound > 0) %>% 
  group_by(iyear) %>% 
  summarise(Killed = sum(nkill), Wounded = sum(nwound)) %>% 
  ungroup() %>% 
ggplot(aes(x = iyear)) +
  geom_line(aes(y = Killed, colour = "Killed")) + 
  geom_point(aes(y = Killed, colour = "Killed")) + 
  geom_line(aes(y = Wounded, colour = "Wounded")) + 
  geom_point(aes(y = Wounded, colour = "Wounded")) + 
  scale_x_continuous(breaks = terrorists_attack$iyear) +
  labs(title = "", x = "Year", y = "", colour = "") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45,
                       hjust = 1,
                       vjust = 1))
```

<br> 
  
#### Number Attacks Based on Governate
```{r eval = TRUE, echo = FALSE}
ggplot(terrorists_attack, aes(x = provstate, fill = provstate)) +
  geom_bar() +
  labs(title = "", x = "Governate", y = "", fill = "") +
  geom_text(stat = "count", aes(label = ..count..), 
            vjust = -0.2, size = 3) +
  theme_bw() +
  theme(legend.position = "none")
```

Since most of attacks occurred in __Capital, Central and Northen__. We will
display the attacks for these governates per city.

<br>

#### Attacks per city based on Capital, Central and Northen governates
```{r eval = TRUE, echo = FALSE}
terrorists_attack %>% 
  filter(provstate %in% c("Capital", "Central", "Northern")) %>% 
ggplot(., aes(x = city, fill = provstate)) +
  geom_bar() +
  labs(title = "", x = "City", y = "", fill = "Governate") +
  coord_flip() +
  geom_text(stat = "count", aes(label = ..count..), 
            vjust = .3, hjust = -.5, size = 3) +
  theme_bw()

```

<br>

#### Attack type and target type
```{r eval = TRUE, echo = FALSE}
ggplot(terrorists_attack, aes(x = targtype1_txt, fill = attacktype1_txt)) +
  geom_bar(position = "dodge") +
  labs(title = "", x = "Target type", y = "", fill = "Attack Type") +
  scale_y_continuous(breaks = seq(0, 50, by = 5)) +
  theme_bw() +
  theme(legend.position = "top", 
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

<br>

#### Number of Killed/Wounded by group attacker
```{r eval = TRUE, echo = FALSE}
terrorists_attack %>% 
  filter(nkill > 0 | nwound > 0) %>% 
  group_by(gname) %>% 
  summarise(killed = sum(nkill), wounded = sum(nwound)) %>% 
  gather(type, value, -gname) %>% 
ggplot(., aes(x = reorder(gname, value), y = value, fill = type)) +
  geom_col(position = "dodge") +
  labs(title = "", x = "", y = "", fill = "") +
  geom_text(aes(label = value), position = position_dodge(width = 1), 
            hjust = -0.1, vjust = 0.3, size = 3) +
  coord_flip() +
  theme_bw() +
  theme(legend.position = "top")
```
<br>

#### Locations of attacks
```{r eval = TRUE, echo = FALSE}
attacks_locations <- terrorists_attack %>% 
  filter(!is.na(terrorists_attack$longitude), eventid != 199112310005)
leaflet(attacks_locations) %>% 
  addTiles(urlTemplate = "http://{s}.tile.openstreetmap.org/{z}/{x}/{y}.png") %>% 
  addMarkers(lat = attacks_locations$latitude, 
             lng = attacks_locations$longitude,
             clusterOptions = markerClusterOptions(),
             popup = paste("<strong>Date: </strong>", attacks_locations$iday,"/",attacks_locations$imonth,"/", attacks_locations$iyear,
                          "<br><br><strong>Main weapon: </strong>", ifelse(!is.na(attacks_locations$weaptype1_txt), attacks_locations$weaptype1_txt, "NA"),
                          "<br><strong>Killed: </strong>", attacks_locations$nkill,
                          "<br><strong>Wounded: </strong>", attacks_locations$nwound,
                          "<br><strong>Summay: </strong>", ifelse(!is.na(attacks_locations$summary), attacks_locations$summary, "NA")
                          
             ))
```
<br>

_Please note that 9 locations had been removed due to missing
longitude/latitude or wrong data._